---
title: "Practical Machine Learning Assignment"
author: "fionawschan"
date: "7/27/2020"
output: html_document
---

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Data Loading and Processing
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r, include=FALSE}
# Loading libraries
library(e1071)
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
library(gbm)
set.seed(12345)
```

```{r}
# Download data from URL
training <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
testing <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))

# Create a partition with the training dataset
inTrain <- createDataPartition(training$classe, p=0.7, list=FALSE)
Trainset <- training[inTrain, ]
Testset <- training[-inTrain, ]
dim(Trainset)
```

```{r} 
dim(Testset)
```

``` {r}
# Remove variables with Nearly Zero Variance
NZV <- nearZeroVar(Trainset)
Trainset <- Trainset[, -NZV]
Testset <- Testset[, -NZV]
dim(Trainset)
```

```{r}
dim(Testset)
```


```{r}
AllNA <- sapply(Trainset, function(x) mean(is.na(x))) > 0.95
Trainset <- Trainset[, AllNA==FALSE]
Testset <- Testset[, AllNA==FALSE]
dim(Trainset)
```

```{r}
dim(Testset)
```

```{r} 
# Remove identification only variable (Columns 1 to 5)
Trainset <- Trainset[, -(1:5)]
Testset <- Testset[, -(1:5)]
dim(Trainset)
```

```{r}
dim(Testset)
```

## Correlation Analysis

Correlation among variables is analysed before proceeding to the modelling procedures.
```{r}
corMatrix <- cor(Trainset[, -54])
corrplot(corMatrix, order = 'FPC', method='color', type ='lower', tl.cex = 0.8, tl.col = rgb(0,0,0))
```

## Prediction Model Building

The Decision Tree method applied to model the regressions. 

#### Decision Tree

```{r}
# Model fit
set.seed(12345)
modFitDecTree <- rpart(Trainset$classe ~ ., data=Trainset, method = "class")
fancyRpartPlot(modFitDecTree)
```

```{r}
# Prediction on Test dataset
predictDecTree <- predict(modFitDecTree, newdata = Testset, type = "class")
confMatDecTree <- confusionMatrix(predictDecTree, Testset$classe)
confMatDecTree
```

```{r}
# plot matrix results
plot(confMatDecTree$table, col = confMatDecTree$byClass,
     main = paste("Decision Tree - Accuracy = ",
                  round(confMatDecTree$overall['Accuracy'],4)))
```

## Conclusion

The accuracy of the Decision Tree modelling method is 0.7342. 
